---
title: "AISanalyze: User guide"
author: "Rémi Pigeault"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AISanalyze: User guide}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(AISanalyze)
library(dplyr)
library(lubridate)
```

# Introduction

AISanalyze provides a workflow to analyse Automatic Identification System (AIS)
data, including:
  
- estimating vessel travel distance, time and speed;
- correcting GPS errors and delays;
- identifying AIS stations and aircraft;
- interpolating vessel positions;
- extracting vessels around target locations;
- estimating vessel characteristics.

This vignette illustrates a typical workflow.

# Example data

```{r}
data("ais")
data("point_to_extract")
```

Convert timestamps to Unix time.

```{r}
ais <- ais %>%
  mutate(timestamp = as.numeric(ymd_hms(datetime)))

point_to_extract <- point_to_extract %>%
  mutate(timestamp = as.numeric(ymd_hm(datetime)))
```

# Estimate travelled distance and speed

```{r}
ais <- AIStravel(ais_data = ais)
```

Three variables are added:
  
- `distance_travelled`
- `time_travelled`
- `speed_kmh`

# Identify stations and aircraft

```{r}
ais <- AISidentify_stations_aircraft(ais_data = ais)
```

Two logical variables are added:
  
- `station`
- `high_speed`

# Correct GPS errors

```{r}
ais <- AIScorrect_speed(ais_data = ais)
```

This step corrects unrealistic speeds caused by GPS errors or transmission delays.

# Interpolate vessel positions

The example below interpolates vessel positions every 60 seconds.

```{r eval=FALSE}
ais_interp <- AISinterpolate(
  ais_data = ais,
  type_interpolation = "maximum_time_interval",
  maximum_time_interval = list(
    maximum_gap_seconds = 60
  )
)
```

The `datetime` column can be updated from the interpolated timestamps:

```{r eval=FALSE}
ais_interp$datetime <- lubridate::as_datetime(timestamp)
```

Alternatively, interpolation can be performed at exact timestamps. Target
locations and a search radius (m) can be specified to limit interpolation to
the area of interest and reduce computation time. The `datetime` column can
then be updated from the new `timestamp`.

```{r eval=FALSE}
ais_interp <- AISinterpolate(
  ais_data = ais %>% AIStravel(),
  type_interpolation = "exact_timestamp",
  exact_timestamp = list(
    timestamp_to_interpolate = point_to_extract$timestamp,
    locations_of_interest = point_to_extract[c("lon", "lat")],
    radius = 200000
  )
)

ais_interp$datetime <- lubridate::as_datetime(timestamp)
```

# Extract nearby vessels

Extract vessel positions within 50 km and ±5 minutes of the target locations and
timestamps (`point_to_extract`).

```{r eval=FALSE}
AISextract(
  data = point_to_extract,
  ais_data = ais,
  return_all_vessel_locations = TRUE,
  search_into_radius_m = 50000,
  interval_time_before = 300,
  interval_time_after = 300
)
```

Set `return_all_vessel_locations = FALSE` to return only the vessel position
closest in time to each target timestamp.

# Estimate vessel characteristics

```{r eval=FALSE}
infos <- AISinfos(ais)

summary_values <- infos$summary
estimated_values <- infos$estimated_values
```

This function estimates the most likely vessel characteristics for each MMSI,
including ship type, dimensions, draught, IMO number, and name.
`summary_values` summarises all values found in the AIS data, whereas
`estimated_values` contains the estimated characteristic for each vessel.

# Workflow summary

The recommended workflow is:
  
  ```text
AIS data
│
▼
AIStravel()
│
▼
AISidentify_stations_aircraft() (optional)
│
▼
AIScorrect_speed() (optional)
│
▼
AISinterpolate()   (optional)
│
▼
AISextract()
│
▼
AISinfos()         (optional)
```
